
# Results

## Descriptives

### Covariates: SES, BMIZ at Age 13, ED Cognitions at age 14

Descriptives for continuous variables are provided in \@ref(tab:continuous)

```{r continuous}
load('data/demo_tables.Rdata')
knitr::kable(cont_vars, caption = 'Continuous Variable Descriptives')
```

Descriptives for ordinal variables (Parent SES and Fear of Weight Gain) are provided in Table \@ref(tab:pses) and Table \@ref(tab:fearwt)

```{r pses}
knitr::kable(pses_table, caption = 'Parent Highest Occupational Class - Responses Frequencies')
```
```{r fearwt}
knitr::kable(fear_wt_table, caption = 'Fear of Weight Gain at Age 14 - Responses Frequencies')
```





### Exercise for Weight Loss

Figure \@ref(fig:freqfig) shows exercise status at ages 14-24 years in girls and boys, as a percentage of the total number of boys and girls who reported exercise data at the timepoint.

```{r freqfig, fig.caption = 'Frequency of Exercise Groups Across Age'}
exercise_group_data <- frq_table_by_age_sex(demo_ed.lf, exercise_group)

exercise_status_plot_1 <- ggplot(data = exercise_group_data, 
    aes(x = sex, y = pct, fill = exercise_group, label = sprintf('%0.1f%%', pct*100))) + #set the aesthetic features, including the labels at one decimal place
    geom_bar(position = position_stack(reverse = TRUE), stat = 'identity') + #add column graph, with male and female separated
    scale_y_continuous(labels = scales::percent, limits = c(0,1))+ #set the lower and upper limits of the y axis ( 0-100 percent); 
    theme(legend.title = element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), text = element_text(size = 15))+ #make legend text a bit bigger
    labs(x = element_blank(), y = 'Percent (within gender)', title = 'Exercise Groups Across Age') +
    facet_wrap(~age) + 
    theme_classic()+
    theme(legend.title = element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), text = element_text(size = 15)) + #make legend text a bit bigger 
   scale_fill_manual(values = wes_palette(name = 'GrandBudapest1', n = 3 ))
exercise_status_plot_1
```

Overall, it appears that percentage of males and female in the three exercise groups is relatively stable from ages 14-18 years, with a possible increase in exercise for weight loss at age 18 among girls, and a more substantive increase in exercise for weight loss among men at age 24 along with an increase in driven exercise across gender at age 24.


